### 2022-11-02 ArticleAggregator_v05_pure.qmd
Initial upload

---
title: "Peripheral Edema measurement methods (by ArticleAggregator_v03)"
author: "ReCodeR"
format: html
server: shiny
editor: 
  markdown: 
    wrap: 72
---

## Sources of Articles

You should use:

-   Scopus
-   Web of Science
    -   [wosr](https://cran.r-project.org/web/packages/wosr/wosr.pdf)
-   Pubmed
-   EMBASE (for posters)
-   Scholar (including patents)

## Code
Use approach from [vignette](http://cran.nexr.com/web/packages/easyPubMed/vignettes/easyPM_vignette_html.html)

```{r}
#| label: Install packages
install.packages("rvest")
install.packages("dplyr")
install.packages("httr")
install.packages("writexl")
install.packages("easyPubMed")

```

```{r}
#| label: Load packages
library("rvest")
library("dplyr")
library("httr")
library("stringr")
library("glue")
library("writexl")
library("purrr")
library("readxl")
library("xml2")
library("easyPubMed")
```

```{r}
#| label: Environment Setup
final_col_names <- c("Title", "Author", "Year", "FTlink", "Src", "Info", "PubDate", "PubType", "Abstract", "CT",      "PMID", "PUI", "DOI", "EBlink", "OpenLink") 
term <- "'peripheral edema' measurement"
aa_path <- r"(C:\Users\I0172484\Documents\01 - Projects\LevoAmlo 2022 Cecko\Edema Study Oct2022\Method Literature Review\)"
setwd(aa_path)
```

## EMBASE Manual export

```{r}
#| label: Load file exported from EMBASE with standard columns
eb_std_col <- read_xlsx(glue("{aa_path}Edema_Methods_eb_exp.xlsx"))
eb_std_col_names <- c("Title", "AuthorAll", "Year", "PubDate", "PubType", "Abstract", "CT", "PMID", "PUI", "DOI", "FTlink", "EBlink", "OpenLink")
names(eb_std_col) <- eb_std_col_names

eb <- eb_std_col %>% 
  mutate(Author=str_extract(AuthorAll, pat="\\w+(?=\\s)")) %>% 
  select(-AuthorAll) %>% 
  mutate(Info=NA) %>% 
  mutate(Src="EMBASE") %>% 
  select(Title, Author, Year, FTlink, Src, Info, everything())

# write_xlsx(eb, glue("{aa_path}eb_to_merge.xlsx"))
# Correct for Van Der Lee manually

```

## Google Scholar Using rvest

```{r}
#| label: Load Google Scholar page
gs_term <- str_replace_all(term, " ", "+")
my_uri <- glue("https://scholar.google.com/scholar?hl=en&as_sdt=0%2C5&q={gs_term}")
gs <- read_html(my_uri)
n_hits <- gs %>% html_elements("#gs_ab_md .gs_ab_mdw") %>% html_text() %>% str_extract("\\d+") %>% as.integer()
n_pg <- n_hits %/% 10 + 1
# n_pg <- gs %>% html_elements("#gs_n a") %>% html_text() %>% as.integer() %>% na.omit() %>% max()

res_df <- tibble(
  title = character(),
  info = character(),
  ref = character()
)

for (i in 0:n_pg){
  curr_pg <- i*10
  print(curr_pg)
  my_uri_pg <- glue("https://scholar.google.com/scholar?start={curr_pg}&q={gs_term}&hl=en&as_sdt=0,5&as_vis=1")
  curr_pg <- read_html(my_uri_pg)
  art_title <- curr_pg %>% html_elements("h3 a") %>% html_text()
  art_info <- curr_pg %>% html_elements(".gs_a") %>% html_text()
  art_ref <- curr_pg %>% html_elements(".gs_rt") %>% html_elements("a") %>% html_attr("href")

  curr_df <- data.frame(title=art_title, info=art_info, ref=art_ref)
  res_df <- bind_rows(res_df, curr_df)
}

# art_title <- gs %>% html_elements("h3 a") %>% html_text()
# art_info <- gs %>% html_elements(".gs_a") %>% html_text()
# art_ref <- gs %>% html_elements(".gs_rt") %>% html_elements("a") %>% html_attr("href")
# output_df <- data.frame(title=art_title, info=art_info, ref=art_ref)
# write_xlsx(res_df, glue("{aa_path}gs_raw.xlsx"))
```

```{r}
#| label: Split content of column "info"
# res_df <- read_xlsx(glue("{aa_path}gs_raw.xlsx"))
sep_df <- res_df %>% 
  mutate(year=str_extract(info, pat="\\d+")) %>% 
  mutate(author=str_extract(info, pat="(?<=[A-Z]\\s([A-Z]\\s)?)\\w+(?=\\,\\s|\\s-)")) %>% 
  select(title, author, year, info, ref)

# rename colums to be compatible with EMBASE and fill missing columns
names(sep_df) <- c("Title", "Author", "Year", "Info", "FTlink")

```

```{r}
#| label: Expand Google Search (gs) columns to fit with EMBASE
gs <- sep_df %>% 
  mutate(PubDate = NA) %>% 
  mutate(PubType = NA) %>%
  mutate(Abstract = NA) %>% 
  mutate(CT = NA) %>%
  mutate(PMID = NA) %>% 
  mutate(PUI = NA) %>%
  mutate(DOI = NA) %>% 
  mutate(EBlink = NA) %>%
  mutate(OpenLink = NA) %>% 
  mutate(Src = "GoogleScholar") %>% 
  select(Title, Author, Year, FTlink, Src, Info, everything())

write_xlsx(gs, glue("{aa_path}gs_to_merge.xlsx"))
## Manually input missing values ;)
```
## Pubmed

You can get full-text link using Elink method of NIH e-utilities tool

```{r}
#| label: PubMed connection
api_key <- "0b1204719f054732c66e608d9d3f2c749c08"
pm_query <- glue('peripheral edema measurement')
my_entrez_id <- get_pubmed_ids(pm_query, api_key=api_key)
# my_abstracts_txt <- fetch_pubmed_data(my_entrez_id, format = "abstract")
# print(my_abstracts_txt[1:45])

my_abstracts_xml <- fetch_pubmed_data(pubmed_id_list = my_entrez_id)
print(paste0("Query retireved ",my_entrez_id$Count," records."))
# print(my_abstracts_xml)
# write(my_abstracts_xml, file = "PubMed_records_by_write.xml")
# recs <- read_xml(my_abstracts_xml) #  From DB connection
recs <- read_xml(paste0(aa_path, "PubMed_records_by_write.xml")) # From file

# USE THIS  TO EXTRACT ALL FIELDS AND CHARACTER VECTOR AND JOIN TO DATAFRAME

xml_abstract_n <- recs %>% xml_find_all(".//Abstract") %>% length()
xml_pmid_n <- recs %>% xml_find_all(".//MedlineCitation/PMID | .//BookDocument/PMID") %>% length()
xml_year_n <- recs %>% xml_find_all(".//PubDate/Year | .//PubDate/MedlineDate") %>% length()
xml_author_n <- recs %>% xml_find_all(".//AuthorList") %>% xml_find_first(".//Author/LastName") %>% length()
xml_title_n <- recs %>% xml_find_all(".//ArticleTitle") %>% length()

# CONTINUE HERE ----
# Some articles have Either EIdType OR IdType OR BOTH.I need IF to remove those with both
xml_doi_n <- recs %>% xml_find_all(".//Article/ELocationID[@EIdType='doi']") %>% length() # N201
xml_doi_n <- recs %>% xml_find_all(".//PubmedData/ArticleIdList/ArticleId[@IdType='doi'] | .//Article/ELocationID[@EIdType='doi']") %>% length() # N528
# WRONG if
# inspire here: https://stackoverflow.com/questions/30604107/r-conditional-evaluation-when-using-the-pipe-operator
xml_doi_n <- recs %>% {if(xml_find_all(".//PubmedData/ArticleIdList/ArticleId[@IdType='doi']") . else xml_find_all(".//Article/ELocationID[@EIdType='doi']")} %>% length()

art <- ".//PubmedArticle | .//PubmedBookArticle"
e_doi <- ".//Article/ELocationID[@EIdType='doi']"
p_doi <- ".//PubmedData/ArticleIdList/ArticleId[@IdType='doi']"
n_art <- recs %>% xml_find_all(art)
xml_doi_n <- recs %>% xml_find_all(art) %>%
  map(if(is_empty(xml_find_first(p_doi)))
  xml_find_first(e_doi) else
  xml_find_first(p_doi)) %>% 
  length()

xml_doi_n <- recs %>% xml_find_all(art) %>%
  map(., ~ifelse(is_empty(xml_find_first(.x, p_doi)), 
  xml_find_first(e_doi),
  xml_find_first(p_doi))) %>% 
  length()

{if(is_empty(xml_find_all(".//PubmedData/ArticleIdList/ArticleId[@IdType='doi']"))
                          . else xml_find_all(".//Article/ELocationID[@EIdType='doi']")} %>% length()
# LOOP: take each article. IF "have both" take "IdType" only. Do for each node and append doi to resulting vector. Use for loop.

# Count of doi should be 372 BUT is 327 :) Correction needed.
xml_pubtype_n <- recs %>% xml_find_all(".//PublicationTypeList | .//BookDocument") %>% xml_find_first(".//PublicationType") %>% length()
xml_nodes_count <- data.frame(xml_abstract_n, xml_pmid_n, xml_year_n, xml_author_n, xml_title_n, xml_doi_n, xml_pubtype_n)
print(xml_nodes_count)
rm(list = ls(pattern='_n$'))

### Split xml by items to data.frame rows ----
library(dplyr)
library(xml2)
library(purrr)
library(tidyr)
xml <- xml2::read_xml("input/example.xml")
rows <- xml %>% xml_find_all("//xmlsubsubnode")
rows_df <- data_frame(node = rows) %>%
  mutate(node_id = node %>% map(~ xml_find_first(., "ancestor::xmlnode")) %>% map(~ xml_attr(., "node-id"))) %>%
  mutate(subnode_id = node %>% map(~ xml_parent(.)) %>% map(~ xml_attr(., "subnode-id"))) %>%
  mutate(text = node %>% map(~ xml_text(.))) %>%
  select(-node)
# https://stackoverflow.com/questions/49253021/how-to-extract-xml-attr-and-xml-text-on-different-levels-with-xml2-and-purrr

xml_articles <- recs %>% xml_find_all(".//PublicationTypeList | .//BookDocument")
xml_articles[[1]][["doc"]]
df_articles <- tibble(item = xml_articles)
ls_articles <- as.list(xml_articles)

### 5 Cases for DOI ----
xml_5_doi <- recs %>%
  xml_find_all(".//Article[@PubModel='Print']/../../PubmedData/ArticleIdList/ArticleID[@IdType='doi']") %>%
  length()

### THIS WORKS ----
xml_5_doi_tested <- recs %>%
  xml_find_all("//Article[@PubModel='Print']/ancestor::PubmedArticle/PubmedData/ArticleIdList/ArticleId[@IdType='doi']") %>%
  xml_text()

xml_5_doi <- recs %>%
  xml_find_all("//Article[@PubModel='Print']/ancestor::PubmedArticle/PubmedData/ArticleIdList/ArticleId[@IdType='doi'] |
                //Article[@PubModel='Print-Electronic']/ancestor::PubmedArticle/PubmedData/ArticleIdList/ArticleId[@IdType='doi'] |
                //Article[@PubModel='Electronic']/ancestor::PubmedArticle/PubmedData/ArticleIdList/ArticleId[@IdType='doi'] |
                //Article[@PubModel='Electronic-eCollection']/ancestor::PubmedArticle//ELocationID[@EIdType='doi']
               ") %>%
  length()

### CODE below generates 327 instead of 371 Captures ----
xml_5_doi <- recs %>%
  xml_find_all("//Article[@PubModel='Print']/ancestor::PubmedArticle/PubmedData/ArticleIdList/ArticleId[@IdType='doi'] |
  //Article[@PubModel='Print-Electronic']/ancestor::PubmedArticle/PubmedData/ArticleIdList/ArticleId[@IdType='doi'] |
  //Article[@PubModel='Electronic']/ancestor::PubmedArticle/PubmedData/ArticleIdList/ArticleId[@IdType='doi'] |
  //Article[@PubModel='Electronic-eCollection']/ancestor::PubmedArticle//ELocationID[@EIdType='doi']
  ") %>%
  length()

# Analysis of problematic PubModel ----
recs %>% xml_find_all("//Article[@PubModel]='Electronic'") %>% length() # Correct, 371 because one is PubmedBookArticle

recs %>% xml_find_all("//PubmedArticle") %>% length() # Correct, 371 because one is PubmedBookArticle

recs %>% xml_find_all("//Article[@PubModel='Print']") %>% length() # 207
recs %>% xml_find_all("//Article[@PubModel='Print']/ancestor::PubmedArticle/PubmedData/ArticleIdList/ArticleId[@IdType='doi']") %>% length() #165

recs %>% xml_find_all("//Article[@PubModel='Print-Electronic']") %>% length() # 138
recs %>% xml_find_all("//Article[@PubModel='Print-Electronic']/ancestor::PubmedArticle/PubmedData/ArticleIdList/ArticleId[@IdType='doi']") %>% length() # 137

recs %>% xml_find_all("//Article[@PubModel='Electronic']") %>% length() # 21
recs %>% xml_find_all("//Article[@PubModel='Electronic']/ancestor::PubmedArticle/PubmedData/ArticleIdList/ArticleId[@IdType='doi']") %>% length() # 20

recs %>% xml_find_all("//Article[@PubModel='Electronic-eCollection']") %>% length() # 5
recs %>% xml_find_all("//Article[@PubModel='Electronic-eCollection']/ancestor::PubmedArticle/PubmedData/ArticleIdList/ArticleId[@IdType='doi']") %>% length() #5

PubmedArticle_n <- recs %>% xml_find_all("/PubmedArticleSet/PubmedArticle") %>% length() # Is correctly 371; PubmedArticle[369] missing; does not have ANY DOI
PubmedArticle_369 <- recs %>% xml_find_all("/PubmedArticleSet/PubmedArticle[369]//ArticleTitle") %>% xml_text()

models <- recs %>% xml_find_all("//Article[@PubModel]") %>% xml_attr("PubModel") %>% as.factor() %>% levels()
models

# Find which nodes are missing in 327 nodes. ----
node_path_n <- recs %>% xml_find_all("//*[@*='doi']") %>% length()
node_path <- recs %>% xml_find_all("//*[@*='doi']") %>% xml_path()
node_path
node_mode <- recs %>% xml_find_all("//*[@*='doi']/ancestor::Article[@PubModel]") %>% length()
write.csv(node_path, glue("{aa_path}node_path.csv"))

# xml_5_doi <- recs %>% 
#   xml_find_all(".//Article[@PubModel='Print']/ancestor::PubmedArticle") %>% xml_text()
#   xml_find_all(".//ArticleID[@IdType='doi']")
#   length()

xml_abstract <- recs %>% xml_find_all(".//Abstract") %>% xml_text()
xml_pmid <- recs %>% xml_find_all(".//MedlineCitation/PMID | .//BookDocument/PMID") %>% xml_text()
xml_year <- recs %>% xml_find_all(".//PubDate/Year | .//PubDate/MedlineDate") %>% xml_text()
xml_author <- recs %>% xml_find_all(".//AuthorList") %>% xml_find_first(".//Author/LastName") %>% xml_text()
xml_title <- recs %>% xml_find_all(".//ArticleTitle") %>% xml_text()
xml_doi <- recs %>% xml_find_all(".//PubmedData/ArticleIdList/ArticleId[@IdType='doi']") %>% xml_text()
xml_pubtype <- recs %>% xml_find_all(".//PublicationTypeList | .//BookDocument") %>% xml_find_first(".//PublicationType") %>% xml_text()

pm_src <- data.frame(xml_title, 
                 xml_author,
                 xml_year,
                 xml_doi,
                 xml_pmid,
                 xml_abstract,
                 xml_pubtype)

pm <- pm_src %>% 
  mutate(FTlink = paste0("https://dx.doi.org/",pm_src$xml_doi)) %>% 
  mutate(CT = NA) %>% 
  mutate(PUI = NA) %>%
  mutate(PubDate = NA) %>%
  mutate(EBlink = NA) %>%
  mutate(OpenLink = NA) %>% 
  mutate(Info = NA) %>% 
  mutate(Src = "PubMed") %>% 
  select(xml_title, 
         xml_author, 
         xml_year, 
         FTlink, 
         Src, 
         Info, 
         PubDate, 
         xml_pubtype, 
         xml_abstract, 
         CT, 
         xml_pmid, 
         PUI, 
         xml_doi,
         EBlink,
         OpenLink)

names(pm) <- final_col_names
write_xlsx(pm, glue("{aa_path}pm_to_merge.xlsx"))
# rm(list = ls(pattern='^xml_'))
# rm(list=ls()[!ls() %in% c("eg")])

```
```{r}
#| label: Interating over PubMed (pm) article nodes
api_key <- "0b1204719f054732c66e608d9d3f2c749c08"
pm_query <- glue('peripheral edema measurement')
my_entrez_id <- get_pubmed_ids(pm_query, api_key=api_key)
# my_abstracts_txt <- fetch_pubmed_data(my_entrez_id, format = "abstract")
# print(my_abstracts_txt[1:45])
my_abstracts_xml <- fetch_pubmed_data(pubmed_id_list = my_entrez_id)
print(paste0("Query retireved ",my_entrez_id$Count," records."))
# print(my_abstracts_xml)
# write(my_abstracts_xml, file = "PubMed_records_by_write.xml")
recs <- read_xml(my_abstracts_xml)

xml_arts <- recs %>% xml_find_all(".//PublicationTypeList | .//BookDocument")
##  CONTINUE HERE ----
art_title <- xml_arts %>% map_dfc(pm_ex(.))
pm_ex <- function(x){ 
  art_title <- x %>% xml_find_first(".//ArticleTitle") %>% xml_text()
  return(art_title)
}
art_title <- xml_arts[[1]] %>% xml_find_first(".//ArticleTitle") %>% xml_text()
print(xml_arts[1])

```
```{r}
#| label: Split items children to tibble / data_frame
xml_in_df <- recs %>% xml_find_all(".//PubmedArticle | .//PubmedBookArticle" ) %>% 
  map(xml_children) %>% 
  map(~as_tibble(t(set_names(xml_text(.), xml_name(.))))) #or map_df
```

```{r}
#| label: Split items to tibble / data_frame
xml_in_df <- recs %>% xml_find_all(".//PubmedArticle | .//PubmedBookArticle" ) %>% 
  map(xml_contents) %>% 
  map(~as_tibble(t(set_names(xml_text(.), xml_name(.))))) #or map_df
```

```{r}
#| label: Merge results from embase (eb), Google Scholar (gs), and PubMed (pm)
eb <- read_xlsx(glue("{aa_path}eb_to_merge.xlsx"))
gs <- read_xlsx(glue("{aa_path}gs_to_merge.xlsx"))
pm <- read_xlsx(glue("{aa_path}pm_to_merge.xlsx"))

egp <- bind_rows(eb, gs, pm) %>% 
  arrange(desc(Year), Author, Src)
# rm(list=ls()[!ls() %in% c("eg")])
# write_xlsx(egp, glue("{aa_path}egp_v02.xlsx"))
```
# ToDo
- correct PubMed functionality
- gs extract surnames that include hyphen
- Introduce SrchIndex
- Add columns: Status, Priority

# ADDITIONAL INFO
### DOI by easyPubmed vignette

```{r}
#| label: PubMed connection
api_key <- "0b1204719f054732c66e608d9d3f2c749c08"
pm_query <- glue('peripheral edema measurement')
my_entrez_id <- get_pubmed_ids(pm_query, api_key=api_key)

# Get articles XML
# my_abstracts_xml <- fetch_pubmed_data(pubmed_id_list = my_entrez_id)
# print(paste0("Query retireved ",my_entrez_id$Count," records."))
# print(my_abstracts_xml)
# write(my_abstracts_xml, file = "PubMed_records_by_write.xml")

# recs <- read_xml(my_abstracts_xml) #  From DB connection
# recs <- read_xml(paste0(aa_path, "PubMed_records_by_write.xml")) # From file

ls_recs <- articles_to_list("PubMed_records_by_write.xml")

# Extract info form the list of articles
pm_df <- ls_recs %>% 
  map_df(~ article_to_df(.x, autofill = TRUE, getAuthors = TRUE, getKeywords = TRUE)) %>% 
  distinct(pmid, .keep_all=TRUE) %>%
  select(title, lastname, year, abstract, pmid, doi)
write_xlsx(pm_df, glue("{aa_path}pm_df_to_merge.xlsx"))

```
### DOI

you can get DOI using CrossRef API, implemented as
https://github.com/ropensci/rcrossref

Find out how many elements of **#gs_n a** are detected. Inspire yourself
in code **WebScrap_Austr_CZ_v12**

```{r}
#| label: Load Google Scholar page

```

```{python}
#| label: Check python installation
import os                   
print(os.environ['path'])  
```

## Retrieve articles from Google Scholar using SerpAPI

Log to the [SerpAPI site](https://serpapi.com/) Install the serpapi
package (package NOT in conda, pip=TRUE must be used):

```{r}
library(reticulate)
py_install("google-search-results", pip=TRUE )
```

Next load the SerpAPI interaction package.

```{python}
from serpapi import GoogleSearch
params = {
  "api_key": "c155eb329cff9b71b6b3cb8edbde7b049ad4cc6ec94a8a53b9a597915e18feb8",
  "engine": "google_scholar",
  "q": "S-lercanidipine",
  "hl": "en"
}
search = GoogleSearch(params)
results = search.get_dict()

```

## Shiny Documents

This Quarto document is made interactive using Shiny. Interactive
documents allow readers to modify parameters and see the results
immediately. Learn more about Shiny interactive documents at
<https://quarto.org/docs/interactive/shiny/>.

## Inputs and Outputs

You can embed Shiny inputs and outputs in your document. Outputs are
automatically updated whenever inputs change. This demonstrates how a
standard R plot can be made interactive:

```{r}
sliderInput("bins", "Number of bins:", 
            min = 1, max = 50, value = 30)
plotOutput("distPlot")
```

```{r}
#| context: server
output$distPlot <- renderPlot({
   x <- faithful[, 2]  # Old Faithful Geyser data
   bins <- seq(min(x), max(x), length.out = input$bins + 1)
   hist(x, breaks = bins, col = 'darkgray', border = 'white',
        xlab = 'Waiting time to next eruption (in mins)',
        main = 'Histogram of waiting times')
})
```