### 2022-11-08 pandas-master/Pandas.qmd
Article Aggreagor PubMed Term added

---
title: "Pandas"
format: html
---

# Code form "Basic Webscraper" by John Watson Rooney
[Source YouTube](https://www.youtube.com/watch?v=vmEHCJofslg)

```{r}
library(reticulate)
py_install("pandas", method="conda" )
py_install("openpyxl", method="conda" )
```

```{python}
#| label: Scrape books from eshop

import pandas as pd
import os

os.getcwd()
df = pd.read_csv("pokemon_data.csv")
# df_x = pd.read_excel("pokemon_data.xlsx")
# df_txt = pd.read_csv("pokemon_data.txt", delimiter='\t')
print(df.head(3))
print(df.tail(3))
print(df.columns)
print(df[['Name', 'Type 1', 'HP']][0:5])

 ```