### 2022-11-08 ArticleAggregator_v05_pure.qmd
ToDo update

---
title: "Peripheral Edema measurement methods (by ArticleAggregator_v03)"
author: "ReCodeR"
format: html
server: shiny
editor: 
  markdown: 
    wrap: 72
---

## Sources of Articles

You should use:

-   Scopus
-   Web of Science
    -   [wosr](https://cran.r-project.org/web/packages/wosr/wosr.pdf)
-   Pubmed
-   EMBASE (for posters)
-   Scholar (including patents)

## Code
Use approach from [vignette](http://cran.nexr.com/web/packages/easyPubMed/vignettes/easyPM_vignette_html.html)

```{r}
#| label: Install packages
install.packages("rvest")
install.packages("dplyr")
install.packages("httr")
install.packages("writexl")
install.packages("easyPubMed")

```

```{r}
#| label: Load packages
library("rvest")
library("dplyr")
library("httr")
library("stringr")
library("glue")
library("writexl")
library("purrr")
library("readxl")
library("xml2")
library("easyPubMed")
```

```{r}
#| label: Environment Setup
final_col_names <- c("Title", "Author", "Year", "FTlink", "Src", "Info", "PubDate", "PubType", "Abstract", "CT",      "PMID", "PUI", "DOI", "EBlink", "OpenLink") 
term <- "peripheral edema measurement"
aa_path <- r"(C:\Users\I0172484\Documents\03 - Scripts\R\WebScrap\ArticleAggregator_data\)"
# setwd(aa_path)
```

## EMBASE Manual export

```{r}
#| label: Load file exported from EMBASE with standard columns
eb_std_col <- read_xlsx(glue("{aa_path}eb_src.xlsx"))
eb_std_col_names <- c("Title", "AuthorAll", "Year", "PubDate", "PubType", "Abstract", "CT", "PMID", "PUI", "DOI", "FTlink", "EBlink", "OpenLink")
names(eb_std_col) <- eb_std_col_names

eb <- eb_std_col %>% 
  mutate(Author=str_extract(AuthorAll, pat="\\w+(?=\\s)")) %>% 
  select(-AuthorAll) %>% 
  mutate(Info=NA) %>% 
  mutate(Src="EMBASE") %>%
  mutate(SrcIndex=row_number()) %>% 
  select(Title, Author, Year, FTlink, Src, Info, everything())

# write_xlsx(eb, glue("{aa_path}eb_to_merge.xlsx"))
# Correct for Van Der Lee manually

```

## Google Scholar Using rvest

```{r}
#| label: Load Google Scholar page
gs_term <- str_replace_all(term, " ", "+")
# my_uri <- glue("https://scholar.google.com/scholar?hl=en&as_sdt=0%2C5&q={gs_term}")

# For Title; including patents; excluding citation
my_uri <- glue("https://scholar.google.com/scholar?hl=en&as_sdt=2007&as_vis=1&q=allintitle%3A+{gs_term}&btnG=")

gs <- read_html(my_uri)
n_hits <- gs %>% html_elements("#gs_ab_md .gs_ab_mdw") %>% html_text() %>% str_extract("\\d+") %>% as.integer()
n_pg <- n_hits %/% 10 + 1
# n_pg <- gs %>% html_elements("#gs_n a") %>% html_text() %>% as.integer() %>% na.omit() %>% max()

res_df <- tibble(
  title = character(),
  info = character(),
  ref = character()
)

for (i in 0:n_pg){
  curr_pg <- i*10
  print(curr_pg)
  my_uri_pg <- glue("https://scholar.google.com/scholar?start={curr_pg}&q={gs_term}&hl=en&as_sdt=0,5&as_vis=1")
  curr_pg <- read_html(my_uri_pg)
  art_title <- curr_pg %>% html_elements("h3 a") %>% html_text()
  art_info <- curr_pg %>% html_elements(".gs_a") %>% html_text()
  art_ref <- curr_pg %>% html_elements(".gs_rt") %>% html_elements("a") %>% html_attr("href")

  curr_df <- data.frame(title=art_title, info=art_info, ref=art_ref)
  res_df <- bind_rows(res_df, curr_df)
}

# art_title <- gs %>% html_elements("h3 a") %>% html_text()
# art_info <- gs %>% html_elements(".gs_a") %>% html_text()
# art_ref <- gs %>% html_elements(".gs_rt") %>% html_elements("a") %>% html_attr("href")
# output_df <- data.frame(title=art_title, info=art_info, ref=art_ref)
# write_xlsx(res_df, glue("{aa_path}gs_src.xlsx"))
```

```{r}
#| label: Split content of column "info"
# res_df <- read_xlsx(glue("{aa_path}gs_raw.xlsx"))
sep_df <- res_df %>% 
  mutate(year=str_extract(info, pat="\\d+")) %>% 
  mutate(author=str_extract(info, pat="(?<=[A-Z]\\s([A-Z]\\s)?)\\w+(?=\\,\\s|\\s-)")) %>% 
  select(title, author, year, info, ref)

# rename colums to be compatible with EMBASE and fill missing columns
names(sep_df) <- c("Title", "Author", "Year", "Info", "FTlink")

```

```{r}
#| label: Expand Google Search (gs) columns to fit with EMBASE
gs <- sep_df %>% 
  mutate(PubDate = NA) %>% 
  mutate(PubType = NA) %>%
  mutate(Abstract = NA) %>% 
  mutate(CT = NA) %>%
  mutate(PMID = NA) %>% 
  mutate(PUI = NA) %>%
  mutate(DOI = NA) %>% 
  mutate(EBlink = NA) %>%
  mutate(OpenLink = NA) %>% 
  mutate(Src = "GoogleScholar") %>% 
  mutate(SrcIndex=row_number()) %>%
  select(Title, Author, Year, FTlink, Src, Info, everything())

# write_xlsx(gs, glue("{aa_path}gs_to_merge.xlsx"))

## Manually input missing values ;)
```
## Pubmed
```{r}
#| label: PubMed connection and article extraction
api_key <- "0b1204719f054732c66e608d9d3f2c749c08"

sl_term <- str_split(term, " ", simplify=TRUE)
pm_query <- str_c(sl_term,"[tiab] ", collapse="")

# pm_query <- "peripheral[tiab] edema[tiab] measurement[tiab]"
# pm_query <- glue("{term}[tiab]")

my_entrez_id <- get_pubmed_ids(pm_query, api_key=api_key)
my_entrez_id$QueryTranslation
my_entrez_id$Count

# Get articles XML
my_abstracts_xml <- fetch_pubmed_data(pubmed_id_list = my_entrez_id, 
                                      retmax=500, 
                                      format="xml", 
                                      encoding="UTF-8")
# print(paste0("Query retireved ",my_entrez_id$Count," records."))
# print(my_abstracts_xml)
write(my_abstracts_xml, file = "PubMed_records_by_write.xml")

# recs <- read_xml(my_abstracts_xml) #  From DB connection
# recs <- read_xml(paste0(aa_path, "PubMed_records_by_write.xml")) # From file

ls_recs <- articles_to_list("PubMed_records_by_write.xml")
# ls_recs <- my_abstracts_xml

# Extract info form the list of articles
pm_df <- ls_recs %>% 
  map_df(~ article_to_df(.x, autofill = TRUE, getAuthors = TRUE, getKeywords = TRUE)) %>% 
  distinct(pmid, .keep_all=TRUE) %>%
  select(title, lastname, year, abstract, pmid, doi)

# Add columns
names(pm_df) <- c("Title", "Author", "Year", "Abstract", "PMID", "DOI")
pm <- pm_df %>% 
  mutate(Info = NA) %>% 
  mutate(PubDate = NA) %>% 
  mutate(PubType = NA) %>%
  mutate(CT = NA) %>%
  mutate(PUI = NA) %>%
  mutate(EBlink = NA) %>%
  mutate(OpenLink = NA) %>%
  mutate(FTlink = paste0("https://dx.doi.org/",pm_df$DOI)) %>% 
  mutate(Src = "PubMed") %>% 
  mutate(SrcIndex=row_number()) %>%
  select(Title, Author, Year, FTlink, Src, Info, everything())
write_xlsx(pm, glue("{aa_path}pm_to_merge.xlsx"))

```
## Merge the results
```{r}
#| label: Merge results from embase (eb), Google Scholar (gs), and PubMed (pm)
eb <- read_xlsx(glue("{aa_path}eb_to_merge.xlsx"))
gs <- read_xlsx(glue("{aa_path}gs_to_merge.xlsx"))
pm <- read_xlsx(glue("{aa_path}pm_to_merge.xlsx"))

egp <- bind_rows(eb, gs, pm) %>%
  mutate(Status=NA) %>% 
  mutate(Priority=NA) %>% 
  arrange(desc(Year), Author, Src) %>% 
  select(Status, Priority, everything())
# rm(list=ls()[!ls() %in% c("eg")])
write_xlsx(egp, glue("{aa_path}egp_v01.xlsx"))
```

# ToDo
- gs extract surnames that include hyphen
- gs with single page outcome
- save final table with search_term name (instead of egp)
- delete variables after writing to xlsx_to_merge
- delete temporary files (to_marge, PubMed.xml etc.)

# ADDITIONAL INFO
### DOI by easyPubmed vignette

```{r}
#| label: PubMed connection
api_key <- "0b1204719f054732c66e608d9d3f2c749c08"
pm_query <- glue('peripheral edema measurement')
my_entrez_id <- get_pubmed_ids(pm_query, api_key=api_key)

# Get articles XML
# my_abstracts_xml <- fetch_pubmed_data(pubmed_id_list = my_entrez_id)
# print(paste0("Query retireved ",my_entrez_id$Count," records."))
# print(my_abstracts_xml)
# write(my_abstracts_xml, file = "PubMed_records_by_write.xml")

# recs <- read_xml(my_abstracts_xml) #  From DB connection
# recs <- read_xml(paste0(aa_path, "PubMed_records_by_write.xml")) # From file

ls_recs <- articles_to_list("PubMed_records_by_write.xml")

# Extract info form the list of articles
pm_df <- ls_recs %>% 
  map_df(~ article_to_df(.x, autofill = TRUE, getAuthors = TRUE, getKeywords = TRUE)) %>% 
  distinct(pmid, .keep_all=TRUE) %>%
  select(title, lastname, year, abstract, pmid, doi)
write_xlsx(pm_df, glue("{aa_path}pm_df_to_merge.xlsx"))

```
### DOI

you can get DOI using CrossRef API, implemented as
https://github.com/ropensci/rcrossref

Find out how many elements of **#gs_n a** are detected. Inspire yourself
in code **WebScrap_Austr_CZ_v12**

```{r}
#| label: Load Google Scholar page

```

```{python}
#| label: Check python installation
import os                   
print(os.environ['path'])  
```

## Retrieve articles from Google Scholar using SerpAPI

Log to the [SerpAPI site](https://serpapi.com/) Install the serpapi
package (package NOT in conda, pip=TRUE must be used):

```{r}
library(reticulate)
py_install("google-search-results", pip=TRUE )
```

Next load the SerpAPI interaction package.

```{python}
from serpapi import GoogleSearch
params = {
  "api_key": "c155eb329cff9b71b6b3cb8edbde7b049ad4cc6ec94a8a53b9a597915e18feb8",
  "engine": "google_scholar",
  "q": "S-lercanidipine",
  "hl": "en"
}
search = GoogleSearch(params)
results = search.get_dict()

```

## Shiny Documents

This Quarto document is made interactive using Shiny. Interactive
documents allow readers to modify parameters and see the results
immediately. Learn more about Shiny interactive documents at
<https://quarto.org/docs/interactive/shiny/>.

## Inputs and Outputs

You can embed Shiny inputs and outputs in your document. Outputs are
automatically updated whenever inputs change. This demonstrates how a
standard R plot can be made interactive:

```{r}
sliderInput("bins", "Number of bins:", 
            min = 1, max = 50, value = 30)
plotOutput("distPlot")
```

```{r}
#| context: server
output$distPlot <- renderPlot({
   x <- faithful[, 2]  # Old Faithful Geyser data
   bins <- seq(min(x), max(x), length.out = input$bins + 1)
   hist(x, breaks = bins, col = 'darkgray', border = 'white',
        xlab = 'Waiting time to next eruption (in mins)',
        main = 'Histogram of waiting times')
})
```