### 2023-02-22 ArticleAggregatorCS.qmd
PubMed fail even with new APIkey

---
title: "refractory erectile dysfunction (by ArticleAggregator_v05)"
author: "ReCodeRa"
format: html
server: shiny
editor: 
  markdown: 
    wrap: 72
---

## Sources of Articles

You should use:

-   Scopus
-   Web of Science
    -   [wosr](https://cran.r-project.org/web/packages/wosr/wosr.pdf)
-   Pubmed
-   EMBASE (for posters)
-   Scholar (including patents)

## Procedure
- EMBASE search: Zentiva Network or VPN, search in "Title, Abstract or Author Keywords"
- Export result in Excel Column Format with following fields: 
    1. Title
    2. Author names
    3. Publication Year
    4. Publication type
    5. Date of publication
    6. Digital Object Identifier (DOI)
    7. Medline PMID
    8. Embase identification number (PUI)
    9. Abstract
    10. Clinical Trial Number
    11. Full text link
    12. Embase link
    13. Open URL Link
- Save to working directory as "eb_src.xlsx"
- Install packages
- Load packages
- Set environment check "term" and "aa_path"
- Decide if use Scholar only for in titles search or full search (enable corresponding url)
- Final Excel table as "egp_v01.xlsx"

## Code
Use approach from [vignette](http://cran.nexr.com/web/packages/easyPubMed/vignettes/easyPM_vignette_html.html)

```{r}
#| label: Install packages
install.packages("rvest")
install.packages("dplyr")
install.packages("httr")
install.packages("readxl")
install.packages("writexl")
install.packages("easyPubMed")

```

```{r}
#| label: Load packages
library("rvest")
library("dplyr")
library("httr")
library("stringr")
library("glue")
library("writexl")
library("purrr")
library("readxl")
library("xml2")
library("easyPubMed")
```

```{r}
#| label: Environment Setup
final_col_names <- c("Title", "Author", "Year", "FTlink", "Src", "Info", "PubDate", "PubType", "Abstract", "CT", "PMID", "PUI", "DOI", "EBlink", "OpenLink") 
term <- "refractory erectile dysfunction"
aa_path <- "/workspaces/WebScrap/ArticleAggregator_data/rami/refrac_ED/"
setwd(aa_path)
```

## EMBASE Manual export

```{r}
#| label: Load file exported from EMBASE with standard columns
eb_std_col <- read_xlsx(glue("{aa_path}eb_src.xlsx"))
eb_std_col_names <- c("Title", "AuthorAll", "Year", "PubDate", "PubType", "Abstract", "CT", "PMID", "PUI", "DOI", "FTlink", "EBlink", "OpenLink")
names(eb_std_col) <- eb_std_col_names

eb <- eb_std_col %>% 
  mutate(Author=str_extract(AuthorAll, pat="\\w+(?=\\s)")) %>% 
  select(-AuthorAll) %>% 
  mutate(Info=NA) %>% 
  mutate(Src="EMBASE") %>%
  mutate(SrcIndex=row_number()) %>% 
  select(Title, Author, Year, FTlink, Src, Info, everything())

write_xlsx(eb, glue("{aa_path}eb_to_merge.xlsx"))
# Correct for Van Der Lee manually

```

## Google Scholar Using rvest

```{r}
#| label: Load Google Scholar page
gs_term <- str_replace_all(term, " ", "+")
my_uri <- glue("https://scholar.google.com/scholar?hl=en&as_sdt=0%2C5&q={gs_term}")

# For Title; including patents; excluding citation
# my_uri <- glue("https://scholar.google.com/scholar?hl=en&as_sdt=2007&as_vis=1&q=allintitle%3A+{gs_term}&btnG=")

gs <- read_html(my_uri)
n_hits <- gs %>% html_elements("#gs_ab_md .gs_ab_mdw") %>% html_text() %>% str_extract("\\d+") %>% as.integer()
n_pg <- n_hits %/% 10 + 1
# n_pg <- gs %>% html_elements("#gs_n a") %>% html_text() %>% as.integer() %>% na.omit() %>% max()

res_df <- tibble(
  title = character(),
  info = character(),
  ref = character()
)

for (i in 0:n_pg){
  curr_pg <- i*10
  print(curr_pg)
  my_uri_pg <- glue("https://scholar.google.com/scholar?start={curr_pg}&q={gs_term}&hl=en&as_sdt=0,5&as_vis=1")
  curr_pg <- read_html(my_uri_pg)
  art_title <- curr_pg %>% html_elements("h3 a") %>% html_text()
  art_info <- curr_pg %>% html_elements(".gs_a") %>% html_text()
  art_ref <- curr_pg %>% html_elements(".gs_rt") %>% html_elements("a") %>% html_attr("href")

  curr_df <- data.frame(title=art_title, info=art_info, ref=art_ref)
  res_df <- bind_rows(res_df, curr_df)
}

# art_title <- gs %>% html_elements("h3 a") %>% html_text()
# art_info <- gs %>% html_elements(".gs_a") %>% html_text()
# art_ref <- gs %>% html_elements(".gs_rt") %>% html_elements("a") %>% html_attr("href")
# output_df <- data.frame(title=art_title, info=art_info, ref=art_ref)
# write_xlsx(res_df, glue("{aa_path}gs_src.xlsx"))
```

```{r}
#| label: Split content of column "info"
# res_df <- read_xlsx(glue("{aa_path}gs_raw.xlsx"))
sep_df <- res_df %>% 
  mutate(year=str_extract(info, pat="\\d+")) %>% 
  mutate(author=str_extract(info, pat="(?<=[A-Z]\\s([A-Z]\\s)?)\\w+(?=\\,\\s|\\s-)")) %>% 
  select(title, author, year, info, ref)

# rename colums to be compatible with EMBASE and fill missing columns
names(sep_df) <- c("Title", "Author", "Year", "Info", "FTlink")

```

```{r}
#| label: Expand Google Search (gs) columns to fit with EMBASE
gs <- sep_df %>% 
  mutate(PubDate = NA) %>% 
  mutate(PubType = NA) %>%
  mutate(Abstract = NA) %>% 
  mutate(CT = NA) %>%
  mutate(PMID = NA) %>% 
  mutate(PUI = NA) %>%
  mutate(DOI = NA) %>% 
  mutate(EBlink = NA) %>%
  mutate(OpenLink = NA) %>% 
  mutate(Src = "GoogleScholar") %>% 
  mutate(SrcIndex=row_number()) %>%
  select(Title, Author, Year, FTlink, Src, Info, everything())

write_xlsx(gs, glue("{aa_path}gs_to_merge.xlsx"))

## Manually input missing values ;)
```
## Pubmed
```{r}
#| label: PubMed connection and article extraction
# api_key <- "0b1204719f054732c66e608d9d3f2c749c08"
api_key <- "7677957ef60aa19aa6c5856e126b696d6408"

sl_term <- str_split(term, " ", simplify=TRUE)
pm_query <- str_c(sl_term,"[tiab] ", collapse="")

# pm_query <- "peripheral[tiab] edema[tiab] measurement[tiab]"
# pm_query <- glue("{term}[tiab]")

my_entrez_id <- get_pubmed_ids(pm_query, api_key=api_key)
my_entrez_id$QueryTranslation
my_entrez_id$Count

# Get articles XML
my_abstracts_xml <- fetch_pubmed_data(pubmed_id_list = my_entrez_id, 
                                      retmax=500, 
                                      format="xml", 
                                      encoding="UTF-8")
# print(paste0("Query retireved ",my_entrez_id$Count," records."))
# print(my_abstracts_xml)
write(my_abstracts_xml, file = "PubMed_records_by_write.xml")

# recs <- read_xml(my_abstracts_xml) #  From DB connection
# recs <- read_xml(paste0(aa_path, "PubMed_records_by_write.xml")) # From file

ls_recs <- articles_to_list("PubMed_records_by_write.xml")
# ls_recs <- my_abstracts_xml

# Extract info form the list of articles
pm_df <- ls_recs %>% 
  map_df(~ article_to_df(.x, autofill = TRUE, getAuthors = TRUE, getKeywords = TRUE)) %>% 
  distinct(pmid, .keep_all=TRUE) %>%
  select(title, lastname, year, abstract, pmid, doi)

# Add columns
names(pm_df) <- c("Title", "Author", "Year", "Abstract", "PMID", "DOI")
pm <- pm_df %>% 
  mutate(Info = NA) %>% 
  mutate(PubDate = NA) %>% 
  mutate(PubType = NA) %>%
  mutate(CT = NA) %>%
  mutate(PUI = NA) %>%
  mutate(EBlink = NA) %>%
  mutate(OpenLink = NA) %>%
  mutate(FTlink = paste0("https://dx.doi.org/",pm_df$DOI)) %>% 
  mutate(Src = "PubMed") %>% 
  mutate(SrcIndex=row_number()) %>%
  select(Title, Author, Year, FTlink, Src, Info, everything())
write_xlsx(pm, glue("{aa_path}pm_to_merge.xlsx"))

```
## Merge the results
```{r}
#| label: Merge results from embase (eb), Google Scholar (gs), and PubMed (pm)
eb <- read_xlsx(glue("{aa_path}eb_to_merge.xlsx"))
gs <- read_xlsx(glue("{aa_path}gs_to_merge.xlsx"))
pm <- read_xlsx(glue("{aa_path}pm_to_merge.xlsx"))

egp <- bind_rows(eb, gs, pm) %>%
  mutate(Status=NA) %>% 
  mutate(Priority=NA) %>% 
  arrange(desc(Year), Author, Src) %>% 
  select(Status, Priority, everything())
# rm(list=ls()[!ls() %in% c("eg")])
write_xlsx(egp, glue("{aa_path}egp_v01.xlsx"))
```

# ToDo
- gs extract surnames that include hyphen
- gs with single page outcome
- save final table with search_term name (instead of egp)
- delete variables after writing to xlsx_to_merge
- delete temporary files (to_marge, PubMed.xml etc.)

# ADDITIONAL INFO
### DOI by easyPubmed vignette

```{r}
#| label: PubMed connection
api_key <- "0b1204719f054732c66e608d9d3f2c749c08"
pm_query <- glue('peripheral edema measurement')
my_entrez_id <- get_pubmed_ids(pm_query, api_key=api_key)

# Get articles XML
# my_abstracts_xml <- fetch_pubmed_data(pubmed_id_list = my_entrez_id)
# print(paste0("Query retireved ",my_entrez_id$Count," records."))
# print(my_abstracts_xml)
# write(my_abstracts_xml, file = "PubMed_records_by_write.xml")

# recs <- read_xml(my_abstracts_xml) #  From DB connection
# recs <- read_xml(paste0(aa_path, "PubMed_records_by_write.xml")) # From file

ls_recs <- articles_to_list("PubMed_records_by_write.xml")

# Extract info form the list of articles
pm_df <- ls_recs %>% 
  map_df(~ article_to_df(.x, autofill = TRUE, getAuthors = TRUE, getKeywords = TRUE)) %>% 
  distinct(pmid, .keep_all=TRUE) %>%
  select(title, lastname, year, abstract, pmid, doi)
write_xlsx(pm_df, glue("{aa_path}pm_df_to_merge.xlsx"))

```
### DOI

you can get DOI using CrossRef API, implemented as
https://github.com/ropensci/rcrossref

Find out how many elements of **#gs_n a** are detected. Inspire yourself
in code **WebScrap_Austr_CZ_v12**

```{r}
#| label: Load Google Scholar page

```

```{python}
#| label: Check python installation
import os                   
print(os.environ['path'])  
```

## Retrieve articles from Google Scholar using SerpAPI

Log to the [SerpAPI site](https://serpapi.com/) Install the serpapi
package (package NOT in conda, pip=TRUE must be used):

```{r}
library(reticulate)
py_install("google-search-results", pip=TRUE )
```

Next load the SerpAPI interaction package.

```{python}
from serpapi import GoogleSearch
params = {
  "api_key": "c155eb329cff9b71b6b3cb8edbde7b049ad4cc6ec94a8a53b9a597915e18feb8",
  "engine": "google_scholar",
  "q": "S-lercanidipine",
  "hl": "en"
}
search = GoogleSearch(params)
results = search.get_dict()

```

## Shiny Documents

This Quarto document is made interactive using Shiny. Interactive
documents allow readers to modify parameters and see the results
immediately. Learn more about Shiny interactive documents at
<https://quarto.org/docs/interactive/shiny/>.

## Inputs and Outputs

You can embed Shiny inputs and outputs in your document. Outputs are
automatically updated whenever inputs change. This demonstrates how a
standard R plot can be made interactive:

```{r}
sliderInput("bins", "Number of bins:", 
            min = 1, max = 50, value = 30)
plotOutput("distPlot")
```

```{r}
#| context: server
output$distPlot <- renderPlot({
   x <- faithful[, 2]  # Old Faithful Geyser data
   bins <- seq(min(x), max(x), length.out = input$bins + 1)
   hist(x, breaks = bins, col = 'darkgray', border = 'white',
        xlab = 'Waiting time to next eruption (in mins)',
        main = 'Histogram of waiting times')
})
```